Massively Parallel Monte Carlo Methods for Discrete Linear and Nonlinear Systems

### 2012-10-24 doc/prelim/presentation/prelim_presentation.tex
adding to presentation

%%---------------------------------------------------------------------------%%
\author{Stuart R. Slattery
  \\ Engineering Physics Department
  \\ University of Wisconsin - Madison
  \\ \href{mailto:sslattery@wisc.edu}{\texttt{sslattery@wisc.edu}}
}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Introduction}
  \begin{itemize}
    \item Predictive modeling and simulation ehances engineering
      capability to yield nuclear system designs that are safer and
      better performing.
    \item Modern work focused on this task leverages multiple physics
      simulation (CASL, NEAMS).
    \item New hardware drives algorithm development (petascale and
      exascale).
    \item Monte Carlo methods have the potential to provide great
      improvements that permit finer simulations and better mapping to
      future hardware.
    \item A set of massively parallel Monte Carlo methods is proposed
      to advance multiple physics simulation on contemporary and
      future leadership class machines.
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Physics-Based Motivation}
  \begin{itemize}
    \item Predictive nuclear reactor analysis enables tighter design
      tolerance for improved thermal performance and efficieny, higher
      fuel burn-up, and high confidence in accident scenario models.
    \item Neutronics, thermal hydraulics, computational fluid
      dynamics, and structural mechanics among other physics
      contribute to these predictive models.
    \item Couplings are complicated and consistent models yield
      nonlinearities in the variables through feedback effects.
    \item  Capturing  these   effects  requires  tremendous  resources
      with $O(\sn{1}{9})$ element meshes and $O(100,000)+$ cores used
      in today's simulations.
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Physics-Based Motivation: DNB}
  \begin{figure}[htpb!]
    \begin{center}
      \scalebox{1}{ \input{dnb_example.pdftex_t} }
    \end{center}
    \caption{\textbf{Multiphysics dependency analysis of departure from
        nucleate boiling.} \textit{A neutronics solution is required to
        compute power generation in the fuel pins, fluid dynamics is
        required to characterize boiling and fluid temperature and
        density, heat transfer is required to compute the fuel and
        cladding temperature, and the nuclear data modified with the
        temperature and density data. Strong coupling among the
        variables creates strong nonlinearities.}}
    \label{fig:dnb_example}
  \end{figure}
\end{frame}

@@

  General stationary methods are formed by splitting the linear
  operator:
  \[
  \ve{A} = \ve{M} - \ve{N}\:.
  \]

@@

  \[
  \ve{e}^{k+1} = \ve{H} \ve{e}^k\:,
  \]

  By diagonalizing $\ve{H}$ to extract its Eigenvalues, we find that 
  \[
  ||\ve{e}^{k}||_2 = \rho(\ve{H})^k ||\ve{e}^0||_2\:,
  \]
  and therefore we bound $\ve{H}$ by $\rho(\ve{H}) < 1$ for convergence.

@@

  \begin{figure}[htpb!]
    \begin{center}
      \scalebox{1.25}{
        \input{orthogonal_residual.pdftex_t} }
    \end{center}
    \caption{\textbf{Orthogonality constraint of the new residual with
        respect to $\mathcal{L}$.} \textit{By projecting $\ve{r}_0$ onto
        the constraint subspace, we minimize the new residual by
        removing those components.}}
    \label{fig:linear_projection_constraint}
  \end{figure}

  From this we note:
  \[
  ||\tilde{\ve{r}}||_2 \leq ||\ve{r}_0||_2,\ \forall \ve{r}_0 \in
  \mathbb{R}^N\:,
  \]
  meaning that the residual of the system will always be
  \textit{minimized} with respect to the constraints.

>>>>

%% colors
\setbeamercolor{boxheadcolor}{fg=white,bg=UWRed}
\setbeamercolor{boxbodycolor}{fg=black,bg=white}

%%---------------------------------------------------------------------------%%
\author{Stuart R. Slattery
  \\ Engineering Physics Department
  \\ University of Wisconsin - Madison
}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Introduction}

  \begin{itemize}
    \item Predictive modeling and simulation ehances engineering
      capability
    \item Modern work focused on this task leverages multiple physics
      simulation (CASL, NEAMS)
    \item New hardware drives algorithm development (petascale and
      exascale)
    \item Monte Carlo methods have the potential to provide great
      improvements that permit finer simulations and better mapping to
      future hardware
    \item A set of massively parallel Monte Carlo methods is proposed
      to advance multiple physics simulation on contemporary and
      future leadership class machines
  \end{itemize}

%%---------------------------------------------------------------------------%%
\begin{frame}{Physics-Based Motivation}

  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Predictive nuclear reactor analysis enables...}
    \begin{itemize}
    \item Tighter design tolerance for improved thermal performance
      and efficieny
    \item Higher fuel burn-up
    \item High confidence in accident scenario models
    \end{itemize}
  \end{beamerboxesrounded}

  \pause 
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
  {Multiple physics simulations are complicated...}
    \begin{itemize}
    \item Neutronics, thermal hydraulics, computational fluid
      dynamics, structural mechanics, and many other physics
    \item Consistent models yield nonlinearities in the variables
      through feedback effects
    \item Tremendous computational resources are required with
      $O(\sn{1}{9})$ element meshes and $O(100,000)+$ cores used in
      today's simulations.
    \end{itemize}
  \end{beamerboxesrounded}

%%---------------------------------------------------------------------------%%
\begin{frame}{Physics-Based Motivation: DNB}
  \begin{figure}[htpb!]
    \begin{center}
      \scalebox{1}{ \input{dnb_example.pdftex_t} }
    \end{center}
    \caption{\textbf{Multiphysics dependency analysis of departure
        from nucleate boiling.} }
    \label{fig:dnb_example}
  \end{figure}
\end{frame}

@@

  \begin{itemize}
  \item General stationary methods are formed by splitting the linear
    operator
  \end{itemize}

  \[
  \ve{A} = \ve{M} - \ve{N}\:.
  \]

@@

  \[
  \ve{e}^{k+1} = \ve{H} \ve{e}^k\:
  \]

  \begin{itemize}
  \item We diagonalize $\ve{H}$ to extract its Eigenvalues
  \end{itemize}

  \[
  ||\ve{e}^{k}||_2 = \rho(\ve{H})^k ||\ve{e}^0||_2\:,
  \]

  \begin{itemize}
  \item We bound $\ve{H}$ by $\rho(\ve{H}) < 1$ for convergence
  \end{itemize}

@@

  \begin{figure}[htpb!]
    \begin{center}
      \scalebox{1.25}{
        \input{orthogonal_residual.pdftex_t} }
    \end{center}
    \caption{\textbf{Orthogonality constraint of the new residual with
        respect to $\mathcal{L}$.} }
  \end{figure}

  \pause
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Minimization Property}

    The residual of the system will always be \textit{minimized} with
    respect to the constraints
    \[
    ||\tilde{\ve{r}}||_2 \leq ||\ve{r}_0||_2,\ \forall \ve{r}_0 \in
    \mathbb{R}^N\:,
    \]
  \end{beamerboxesrounded}
